---
title: "NFL Salary Research"
author: "Austin Miles"
date: "September 9, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ggplot2)
library(readxl)
library(car)
QB_Stats <- read_excel("QBstats.xlsx")
QB_Salary <- read_excel("QBsalaray.xlsx")
for (i in 1: 700){
  for (j in 1: 1340){
    if  (QB_Stats[i, "Player"] == QB_Salary[j, "Player"]){
      if (QB_Stats[i, "Following Year"] == QB_Salary[j, "Year"]){
        QB_Stats[i, "Following Salary"] = QB_Salary[j, "base"]
      }}}}
Following_Salary <- QB_Stats$`Following Salary`
Player_Name <- QB_Stats$Player
Comp <- QB_Stats$Comp
Att <- QB_Stats$Att
Pct <- QB_Stats$Pct
Att_PG <- QB_Stats$`Att/G`
Yards <- QB_Stats$Yds
Avg <- QB_Stats$Avg
Yards_PG <- QB_Stats$`Yds/G`
TD <- QB_Stats$TD
Int <- QB_Stats$Int
First_Downs <- QB_Stats$`1st`
First_Down_Pct <- QB_Stats$`1st%`
Pass_20 <- QB_Stats$`20+`
Pass_40 <- QB_Stats$`40+`
Rating <- QB_Stats$Rate
Year <- QB_Stats$Year
QB_Stats$Log_F.Salary <- log(QB_Stats$`Following Salary`)
```

#Exploratory Analysis

```{r}
fit <- lm(Following_Salary~Comp+Att+Yards+TD+Int+Rating)
scatterplotMatrix(~Following_Salary+Comp+Att+Yards+TD+Int+Rating)
```

```{r}
avPlots(fit)
```

Observation 10 seems to be an outlier.

```{r}
outlierTest(fit)
```

Based on on the calculated P-Values, observation 10, 22, 103, 57 and 39 are significant outliers.
```{r}
summary(fit)
```

The fitted linear model for the Following Salary of a quarterback based on the Completions, Attempts, Yards, Touchdowns, Interceptions and Rating is:

Following Salary =-213998.6+73459.4(Completions)-29649.4(Attempts)-1073(Yards)+15344.3(Touchdowns)+18900.7(Interceptions)+9957.1(Rating).

```{r}
comptest <- lm(Following_Salary~Comp)
atttest <- lm(Following_Salary~Att)
yardstest <- lm(Following_Salary~Yards)
tdtest <- lm(Following_Salary~TD)
inttest <- lm(Following_Salary~Int)
ratingtest <- lm(Following_Salary~Rating)
dfComp <- c("Comp SLR Coefficient", (comptest$coefficients[2]))
dfAtt <- c("Att SLR Coefficient", (atttest$coefficients[2]))
dfYards <- c("Yards SLR Coefficient", (yardstest$coefficients[2]))
dfTD <- c("TD SLR Coefficient", (tdtest$coefficients[2]))
dfInt <- c("Int SLR Coefficient", (inttest$coefficients[2]))
dfRating <- c("Rating SLR Coefficient", (ratingtest$coefficients[2]))
df <- data.frame(dfComp, dfAtt, dfYards, dfTD, dfInt, dfRating)
print.data.frame(df, row.names = FALSE)
```

Completions has a positive relationship with Following Salary, Attempts has a positive relationship with Following Salary, Yards has a positive relationship with Following Salary, Touchdowns has a positive relationship with Following Salary. Surprisingly, Interceptions has a positive relationship with Following Salary and Rating has a positive relationship with Following Salary.

The positive relationship between Interceptions and Following Salary is probably due to the fact that Quarterback's get paid more for how significant they are to the team, and a quarterback with more attempted passes tend to throw more interceptions.

#Tests for Significance of Predictors in Model:

##Completions:
Null Hypothesis $H_{0}: \beta_{1}=0,$ Alternative Hypothesis $H_{A}: \beta_{1} \neq 0$.
```{r}
#T-Value
T=fit$coefficients[2]/13265.2
#Critical Value
CV=qt(0.995, 601)
#Test
abs(T)>CV
```
Since the absolute value of the T value is greater than the critical value at level $\alpha=0.01$, the null hypothesis is rejected and it is determined that $\beta_{1} \neq 0$. Therefore Completions is a significant predictor in the model. 


##Attempts:
Null Hypothesis $H_{0}: \beta_{2}=0,$ Alternative Hypothesis $H_{A}: \beta_{2} \neq 0$.
```{r}
#T-Value
T=fit$coefficients[3]/8574.7
#Critical Value
CV=qt(0.995, 601)
#Test
abs(T)>CV
```
Since the absolute value of the T value is greater than the critical value at level $\alpha=0.01$, the null hypothesis is rejected and it is determined that $\beta_{2} \neq 0$. Therefore Attempts is a significant predictor in the model. 


##Yards:
Null Hypothesis $H_{0}: \beta_{3}=0,$ Alternative Hypothesis $H_{A}: \beta_{3} \neq 0$.
```{r}
#T-Value
T=fit$coefficients[4]/931.5
#Critical Value
CV=qt(0.995,601)
#Test
abs(T)>CV
```
Since the absolute value of the T value is less than the critical value at level $\alpha=0.01$, the null hypothesis is not rejected and it is determined that $\beta_{3} = 0$. Therefore Yards is not a significant predictor in the model. 

##Touchdowns:
Null Hypothesis $H_{0}: \beta_{4}=0,$ Alternative Hypothesis $H_{A}: \beta_{4} \neq 0$.
```{r}
#T-Value
T=fit$coefficients[5]/47055.9
#Critical Value
CV=qt(0.995,601)
#Test
abs(T)>CV
```
Since the absolute value of the T value is less than the critical value at level $\alpha=0.01$, the null hypothesis is not rejected and it is determined that $\beta_{4} = 0$. Therefore Touchdowns is not a significant predictor in the model. 

##Interceptions:
Null Hypothesis $H_{0}: \beta_{5}=0,$ Alternative Hypothesis $H_{A}: \beta_{5} \neq 0$.
```{r}
#T-Value
T=fit$coefficients[6]/55945.8
#Critical Value
CV=qt(0.995,601)
#Test
abs(T)>CV
```
Since the absolute value of the T value is less than the critical value at level $\alpha=0.01$, the null hypothesis is not rejected and it is determined that $\beta_{5} = 0$. Therefore Interceptions is not a significant predictor in the model. 

##Rating:
Null Hypothesis $H_{0}: \beta_{6}=0,$ Alternative Hypothesis $H_{A}: \beta_{6} \neq 0$.
```{r}
#T-Value
T=fit$coefficients[7]/20796.3
#Critical Value
CV=qt(0.995,601)
#Test
abs(T)>CV
```
Since the absolute value of the T value is less than the critical value at level $\alpha=0.01$, the null hypothesis is not rejected and it is determined that $\beta_{6} = 0$. Therefore Rating is not a significant predictor in the model.

# Adding Player Name

The following test will determine if the model is improved enough with the inclusion of the predictor Player Name to warrant the increased complextiy of an added predictor. 
Null Hypothesis $H_0:$ The model without Player Name is sufficient, Alternative Hypothesis $H_A:$ Player Name is a significant addition to the model given the predictors.

```{r}
fullmodel <- lm(Following_Salary~Comp+Att+Yards+TD+Int+Rating+Player_Name)
anova(fit,fullmodel)
```
The value of the test statistic is 1.4582, the null distribution of the test statistic is $F_{1,450}.$ The P-Value is $0.00175}$, which is less than $\alpha=0.05.$ Therefore the null hypothesis is rejected and it is determined that Player Name is a significant addition to the model. 

```{r}
plot(fit, which = 1:3)
plot(fullmodel, which = 1:3)
```
```{r}
c(summary(fit)$r.squared, summary(fullmodel)$r.squared)
```

```{r}
c(AIC(fit), AIC(fullmodel))
```

It is clear that the model with Player Name is a better fit as it has a higher R squared value although the AIC rose slightly.

#Transfroming the Response

```{r}
logtransform <- lm(log(Following_Salary)~Comp+Att+Yards+TD+Int+Rating+Player_Name)
plot(logtransform, which = 1:3)
```

```{r}
summary(logtransform)$r.squared
AIC(logtransform)
```
The logarigthmicallly transformed response improved the model's adherence to the normality of errors assumption and is the best fit out of the models created.

```{r}
New_QB_Stats <- na.omit(QB_Stats)
median(New_QB_Stats$`Following Salary`)
High_QB_Stats <-subset(QB_Stats ,Following_Salary > 1800000)
High_QB_lm <- lm(High_QB_Stats$`Following Salary`~High_QB_Stats$Comp+High_QB_Stats$Att+High_QB_Stats$Yds+High_QB_Stats$TD+High_QB_Stats$Int+High_QB_Stats$Rate+High_QB_Stats$Player)
scatterplotMatrix(~High_QB_Stats$`Following Salary`+High_QB_Stats$Comp+High_QB_Stats$Att+High_QB_Stats$Yds+High_QB_Stats$TD+High_QB_Stats$Int+High_QB_Stats$Rate)
Low_QB_Stats <- subset(QB_Stats, Following_Salary <= 1800000)
Low_QB_lm <- lm(Low_QB_Stats$`Following Salary`~Low_QB_Stats$Comp+Low_QB_Stats$Att+Low_QB_Stats$Yds+Low_QB_Stats$TD+Low_QB_Stats$Int+Low_QB_Stats$Rate+Low_QB_Stats$Player)
scatterplotMatrix(~Low_QB_Stats$`Following Salary`+Low_QB_Stats$Comp+Low_QB_Stats$Att+Low_QB_Stats$Yds+Low_QB_Stats$TD+Low_QB_Stats$Int+Low_QB_Stats$Rate)
plot(High_QB_lm, which = 1:3)
plot(Low_QB_lm, which = 1:3)
summary(High_QB_lm)$r.squared
summary(Low_QB_lm)$r.squared
```
Because of the high R squared value for the model using the dataset of Quarterbacks following salary above the median (of this dataset), we can transform the response of this model to get a stronger fit.

```{r}
logtransform_high <- lm(log(High_QB_Stats$`Following Salary`)~High_QB_Stats$Comp+High_QB_Stats$Att+High_QB_Stats$Yds+High_QB_Stats$TD+High_QB_Stats$Int+High_QB_Stats$Rate+High_QB_Stats$Player)
summary(logtransform_high)$r.squared
```

Since the R squared value is only slightly higher than the log transformed response of the full model which includes all the Quarterbacks in the dataset, we will use that model to estimate a Quarterback's value.

```{r, include=FALSE}
NFL_Ranks <- read_excel("NFL Team Ranks.xlsx")
QB_Stats_Rank <- merge(QB_Stats,NFL_Ranks,by=c("Team","Year"))
QB_Stats_Rank$`OFF. RANK`<- as.numeric(as.character(QB_Stats_Rank$`OFF. RANK`))
```

```{r}
scatterplotMatrix(~QB_Stats_Rank$`OFF. RANK`+QB_Stats_Rank$Comp+QB_Stats_Rank$Att+QB_Stats_Rank$Yds+QB_Stats_Rank$TD+QB_Stats_Rank$Rate)
```
There seems to be some correlation between the Team's offensive rank and the number of Touchdowns, as well as a correlation between the Team's offensive rank and Quarterback rating.

By eliminating Quarterbacks who have thrown less than 10 touchdowns, we can check if the correlation is improved since the visual above includes 2nd or even 3rd string quarterbacks.

```{r}
Top_QB_Rank <- subset(QB_Stats_Rank, QB_Stats_Rank$TD > 10)
scatterplotMatrix(~Top_QB_Rank$`OFF. RANK`+Top_QB_Rank$Comp+Top_QB_Rank$Att+Top_QB_Rank$Yds+Top_QB_Rank$TD+Top_QB_Rank$Rate)
```

```{r}
Top_QB_TD_lm <- lm(Top_QB_Rank$`OFF. RANK`~Top_QB_Rank$TD)
All_QB_TD_lm <- lm(QB_Stats_Rank$`OFF. RANK`~QB_Stats_Rank$TD)
Top_QB_Rate_lm <- lm(Top_QB_Rank$`OFF. RANK`~Top_QB_Rank$Rate)
All_QB_Rate_lm <- lm(QB_Stats_Rank$`OFF. RANK`~QB_Stats_Rank$Rate)
summary(Top_QB_TD_lm)$r.squared
summary(All_QB_TD_lm)$r.squared
summary(Top_QB_Rate_lm)$r.squared
summary(All_QB_Rate_lm)$r.squared
```
As you can see, the offensive rank of the Quarterbacks who throw more than 10 Touchdowns regressed on their team's offensive rank has the highest correlation.

```{r}
summary(Top_QB_Rate_lm)
```

```{r}
ggplot(QB_Stats, aes(x = QB_Stats$TD, y = QB_Stats$Log_F.Salary))+ 
         geom_point(colour = "black", size = .5) + facet_wrap(~QB_Stats$Player)+
  theme(axis.text=element_text(size=5), strip.text = element_text(size = 5, margin = margin()))
```



